### 2014-10-12 paper/craze.tex
Poster craze

\documentclass[paperwidth=120cm,paperheight=90cm,portrait,fontscale=.22,margin=5cm,lmargin=.2cm,rmargin=.2cm]{baposter}

\usepackage{minted}
\definecolor{mintedbg}{rgb}{0.95,0.95,0.95}

\usepackage{upquote}
\AtBeginDocument{%
\def\PYZsq{\textquotesingle}%
}

\definecolor{bordercol}{RGB}{40,40,40}
\definecolor{headercol1}{RGB}{235,218,229}
\definecolor{headercol2}{RGB}{80,80,80}
\definecolor{headerfontcol}{RGB}{0,0,0}
\definecolor{boxcolor}{RGB}{255,235,205}

\renewcommand{\familydefault}{\sfdefault}

\def\mireval{\texttt{mir\char`_eval}}

\usepackage{tabularx}
\usepackage{booktabs}

\begin{document}

\begin{poster}{
  background=plain,
  bgColorOne=white,
  headerheight=0.18\textheight,
  eyecatcher=true,
  columns=1,
  borderColor=bordercol,
  headerColorOne=headercol1,
  headerColorTwo=headercol1,
  headerFontColor=headerfontcol,
  headerfont=\textmd,
  boxColorOne=white,
  linewidth=0.5pt,
  borderColor=black,
  headerborder=none,
  textborder=none,
}
{\includegraphics[height=8em]{labrosa-new-nobg.pdf}}
{\textmd{MIR\char`_EVAL:$\;$A$\;$TRANPARENT$\;$IMPLEMENTATION\\OF$\;$COMMON$\;$MIR$\;$METRICS}}
{Colin Raffel, Brian McFee, Eric J. Humphrey, Justin Salamon, Oriol Nieto, Dawen Liang,\\[.2em] and Daniel P. W. Ellis\\[.2em]}
{\includegraphics[height=6em]{marl-logo.pdf}}

\begin{posterbox}[]{}
\Huge
\mireval{} is a Python module which includes metrics for the following tasks: \textbf{Beat detection, chord estimation, pattern discovery, structural segmentation, melody extraction}, and \textbf{onset detection}.
It's easy to use in Python:
\end{posterbox}

\begin{posterbox}[below=auto]{}

\begin{minted}[bgcolor=boxcolor, frame=single, framerule=0pt, fontsize=\Large]{python}
import mir_eval
# Load in beat annotations
reference_beats = mir_eval.io.load_events('ref_beats.txt')
estimated_beats = mir_eval.io.load_events('est_beats.txt')
# scores will be a dictionary where the key is the metric name and the value is the score achieved
scores = mir_eval.beat.evaluate(reference_beats, estimated_beats)
\end{minted}

\end{posterbox}

\begin{posterbox}[below=auto]{}
\Huge
or without Python, using the included evaluator scripts:
\end{posterbox}

\begin{posterbox}[name=evaluator,below=auto]{}

\begin{minted}[bgcolor=boxcolor, frame=single, framerule=0pt, fontsize=\Large]{bash}
> ./beat_eval.py ref_beats.txt est_beats.txt -o scores.json
> cat scores.json
  {"F-measure": 0.6216216216216, "Cemgil": 0.36267669947376, "Cemgil Best Metric Level": ...
\end{minted}

\end{posterbox}

\begin{posterbox}[below=auto]{}
\Huge
We also quantitatively compared it to MIREX with some intriguing results!
\end{posterbox}

\end{poster}

\end{document}